---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r echo = TRUE, message = FALSE}
library("dplyr")
library("lubridate")
activity <- read.csv(unz("activity.zip", "activity.csv")) %>% 
    tbl_df %>% 
    mutate(date = ymd(date))
``` 


## What is mean total number of steps taken per day?
```{r echo = FALSE, message = FALSE}
library("ggplot2")
totalStepsPerDay <- activity %>% group_by(date) %>% summarise(total = sum(steps))
ggplot(totalStepsPerDay, aes(total)) + geom_histogram() + xlab("Total number of steps taken each day")
``` 

```{r echo = TRUE}
meanStepsPerDay <- as.integer(mean(totalStepsPerDay$total, na.rm = TRUE))
```

Mean steps per day: `r meanStepsPerDay`

```{r echo = TRUE}
medianStepsPerDay <- median(totalStepsPerDay$total, na.rm = TRUE)
```
Median steps per day: `r medianStepsPerDay`


## What is the average daily activity pattern?
```{r echo = TRUE}
meanStepsPerInterval <- activity %>% group_by(interval) %>% summarise(average = mean(steps, na.rm = TRUE)) 
with(meanStepsPerInterval, plot(interval, average, type = "l", xlab = "5 minutes interval", ylab = "Average number of steps", lwd = 2))
```

```{r echo = TRUE}
maxStepInterval <- (meanStepsPerInterval %>% filter(average == max(meanStepsPerInterval$average)))$interval
```

Interval containing maximum number of steps: `r maxStepInterval`


## Imputing missing values
```{r echo = TRUE}
totalMissingValues <- activity %>% filter(is.na(steps)) %>% nrow
```

Total number of missing values in the dataset: `r totalMissingValues`



Fill missing steps with average per step per day:
```{r echo = TRUE}
noNasTotalStepsPerDay <- activity %>% 
    group_by(date) %>% 
    summarise(total = sum(steps)) %>%
    mutate(total = ifelse(is.na(total), mean(total, na.rm = TRUE), total))
```


We can see that in this case, replacing missing values with average significantly
increases the count in the histogram, for x value = mean (as we count expect).

```{r echo = FALSE, message = FALSE}
library("ggplot2")
ggplot(noNasTotalStepsPerDay, aes(total)) + geom_histogram() + xlab("Total number of steps taken each day")
``` 



## Are there differences in activity patterns between weekdays and weekends?
